---
title: "Lab 03 - Nobel laureates"
author: "Saima Arina"
date: "01/31/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

Here is a link to the [lab instructions](https://datascience4psych.github.io/DataScience4Psych/lab03.html).

### Exercise 1

```{r observations and variables}
nrow(nobel)
ncol(nobel)
```

There are 935 observations and there are 26 variables. Each row represents a nobel lauterate individual.

### Exercise 2

```{r, create new df with filters}

nobel_living <- nobel %>% 
  select(firstname, surname, died_date, country, gender, category) %>% 
  filter(is.na(died_date)) %>% 
  filter(!is.na(country))
```


```{r, verifying filters}
nobel %>% 
  count(is.na(died_date))

nobel %>% 
  count(!is.na(country))

nobel %>% 
  count(gender)
```


### Exercise 3

```{r, add category to data frame}

nobel_living <- nobel %>% 
  select(firstname, surname, died_date, country, gender, category) %>% 
  filter(is.na(died_date)) %>% 
  filter(!is.na(country))
```

```{r}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r, filtering categories}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r}
ggplot(data = nobel_living_science,
       mapping = aes(x = country_us, fill = country_us)) +
  geom_bar() +
  facet_wrap(~ category) +
  coord_flip() +
  labs(
    x = "Country at time of prize (USA vs Other)",
    y = "Number of laureates",
    title = "Living Nobel laureates by prize category and whether they were in the US"
)
```
Interpretation

A clear majority of living Nobel laureates in these fields were based in the US when they received their awards. I would say the Buzzfeed headline is supported broadly by the data, although the exact proportions vary depending on the category.

### Exercise 4

```{r, add born_country to data frame}

nobel_living_science <- nobel %>% 
  select(firstname, surname, died_date, country, gender, category, born_country) %>% 
  filter(is.na(died_date)) %>% 
  filter(!is.na(country))
```

```{r}
nobel_living_science <- nobel_living_science %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```


```{r, make new variable}
nobel_living_science <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  )
```

```{r}
nobel_living_science %>% 
  count(born_country_us)
```

...

105 of the winners were born in the U.S.


### Exercise 5

```{r}
ggplot(data = nobel_living_science,
       mapping = aes(x = country_us, fill = born_country_us)) +
  geom_bar() +
  facet_wrap(~ category) +
  coord_flip() +
  labs(
    x = "Country at time of prize (USA vs Other)",
    y = "Number of laureates",
    title = "Living Nobel laureates by prize category and whether they were in the US and origin of birth",
    fill = "Born in US?"
)
```
...
Interpretation:

Yes, the data is supportive. A large amount of laureates who won their prize in the U.S. are classified as "other" for where they were born in every category, indicating that many of the US-based Nobel winners were not born in the U.S.

### Exercise 6

...
