---
title: "Lab 03 - Nobel laureates"
author: "Saima Arina"
date: "01/31/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

Here is a link to the [lab instructions](https://datascience4psych.github.io/DataScience4Psych/lab03.html).

### Exercise 1

```{r observations and variables}
nrow(nobel)
ncol(nobel)
```

There are 935 observations and there are 26 variables. Each row represents a nobel lauterate individual.

### Exercise 2

```{r, create new df with filters}

nobel_living <- nobel %>% 
  select(firstname, surname, died_date, country, gender) %>% 
  filter(is.na(died_date)) %>% 
  filter(!is.na(country))
```


```{r, verifying filters}
nobel %>% 
  count(is.na(died_date))

nobel %>% 
  count(!is.na(country))

nobel %>% 
  count(gender)
```


### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
